Tables & figures for paper reporting increase of home deaths 2020

### 2021-05-06 covid_nrs_place_of_death_visualisations.Rmd
updated cause of death graph and table & made figure 4 from it for the paper

table_deaths_by_cause_2020 %>%
  mutate(across(.cols = matches("ratio|prop"), .fns = ~scales::percent(.x, accuracy=0.1))) %>%
  select(Period = year, "Place of death" = place_of_death, "Cause of death" = cause_of_death, "Annual deaths" = deaths, "Proportion of annual deaths" = proportion_annual_deaths, "Historic average" = annual_deaths_mean, "Historic proportion" = historical_proportion_annual_deaths, "Ratio 2020 to historic" = ratio_annual_deaths_to_historical_mean) %>%
  write_csv(file = paste0(dir_outputs,"/table_causes_of_death_2020.csv"))
```

@@

```{r}
(figure_annual_deaths_by_cause <-
  total_deaths_cause_of_death %>%
  filter(year == 2020 & cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
  # pivot_longer(cols = c(deaths, annual_deaths_mean), names_to = "deaths_type", values_to = "deaths") %>%
  ggplot() +
  # geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths, group = deaths_type, linetype = deaths_type), position = "dodge", colour = "grey10") +
  geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths), colour = "grey10") +
  # geom_point(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  # geom_line(aes(x = cause_of_death, y = annual_deaths_mean), colour = "grey10") +
  scale_fill_viridis_d() +
  facet_wrap(~place_of_death) +
  NULL
)
```

>>>>

## get "canonical" order for causes of death in 2020
order_of_causes_of_death <- table_deaths_by_cause_2020 %>% filter(place_of_death == "All") %>% arrange(desc(deaths)) %>% pull(cause_of_death)

## sort the table by "canonical" order
table_deaths_by_cause_2020 <-
  table_deaths_by_cause_2020 %>%
  group_by(year, place_of_death) %>%
  arrange(cause_of_death = factor(cause_of_death, levels = order_of_causes_of_death), .by_group = TRUE) %>%
  ungroup

table_deaths_by_cause_2020 %>%
  mutate(across(.cols = matches("ratio|prop"), .fns = ~scales::percent(.x, accuracy=0.1))) %>%
  mutate(across(.cols = matches("ratio|^historical"), .fns = ~if_else(condition = cause_of_death=="COVID-19", true = "", false = .x))) %>%
  select(Period = year, "Place of death" = place_of_death, "Cause of death" = cause_of_death, "Annual deaths" = deaths, "Proportion of annual deaths" = proportion_annual_deaths, "Historic average" = annual_deaths_mean, "Historic proportion" = historical_proportion_annual_deaths, "Ratio 2020 to historic" = ratio_annual_deaths_to_historical_mean) %>%
  write_csv(file = paste0(dir_outputs,"/table_causes_of_death_2020.csv"))
```

@@

```{r}
(figure_annual_deaths_by_cause <-
  total_deaths_cause_of_death %>%
  filter(year == 2020 & cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
   mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
   mutate(
     cause_of_death = factor(cause_of_death %>% str_wrap(., width = 15), levels = order_of_causes_of_death %>% str_wrap(., width = 15)),  # needed to allow converting to a number for plotting the line segments; str_wrap breaks the text up into lines
     cause_of_death = fct_rev(cause_of_death)
     ) %>%
  ggplot() +
  geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths), alpha = 0.8, colour = "grey10") +
  geom_segment(aes(x = as.numeric(cause_of_death)-0.45, xend = as.numeric(cause_of_death)+0.45, y = annual_deaths_mean, yend = annual_deaths_mean), size = 0.8, colour = "grey1") +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~place_of_death) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",")) +
  theme(
    legend.position = "none", 
    plot.caption = element_text(size = 10, colour = "gray60"),
    legend.title = element_blank(), 
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
    ) +
  labs(
    x = NULL,
    y = "Deaths",
    subtitle = "Bars are 2020 deaths, line segments are mean annual deaths 2015-2019"
  ) +
  NULL
)

save_output_file(filename = paste0(dir_outputs,"/figure_annual_deaths_by_cause"), extensions = c(".pdf",".svg",".png"), plot = figure_annual_deaths_by_cause, width = 10, height = 3, dpi = 300, units = "in")
```

### 2021-05-06 paper_reporting_death_at_home_increase.Rmd
updated cause of death graph and table & made figure 4 from it for the paper

>>>>

## Causes of death

```{r}
(figure4 <-
  figure_annual_deaths_by_cause +
  scale_fill_grey() +
  labs(x = NULL, y = NULL, title = NULL, subtitle = NULL)
)

save_output_file(filename = paste0(dir_paper,"/figure4"), extensions = c(".pdf",".svg",".png"), plot = figure4, width = 10, height = 3, dpi = 300, units = "in")
save_output_file(filename = paste0(dir_paper,"/figure4_hires"), extensions = c(".png"), plot = figure4, width = 10, height = 3, dpi = 1200, units = "in")
```
